**This is the write-up for the first homework assignment due Tuesday Feb 4.**

1. Use the GEOmetabd package to find all HCV gene expression data using the Illumina platform submitted by an investigator at Yale. This should be done with a single query, showing the title, the GSE accession number, the GPL accession number and the manufacturer and the description of the platform used.

2. Reproduce your above query using the data.table package. Again, try to use a single line of code. [Hint: You first need to convert all db tables to data.table tables].

```{r eval=FALSE}
## eval = FALSE for this chunk under the assumption that the relevant packages are already installed
source("http://bioconductor.org/biocLite.R")
# Install all core packages and update all installed packages
biocLite()
biocLite(c("GEOmetadb", "GEOquery"))
```

```{r}
## eval=TRUE to load the installed package
library(GEOmetadb)
```

```{r eval=FALSE}
## We leave the option to evaluate this as FALSE because it can take a long time to download
getSQLiteFile()
```

```{r}
## IMPORTANT: Working folder must be set where GEOmetadb.sqlite was downloaded
geo_con <- dbConnect(SQLite(),'GEOmetadb.sqlite')
```

```{r}
dbGetQuery(geo_con, "SELECT gse.title, gse.gse, gpl.gpl, gpl.manufacturer, gpl.technology FROM (gpl JOIN gse_gpl ON gpl.gpl=gse_gpl.gpl) j JOIN gse on j.gse=gse.gse WHERE gpl.manufacturer like '%Illumina%' AND gse.contact like '%Yale%' AND (gse.title like '%HCV%' OR gse.title like '%hepatitis C%')")
```

Some additional comments regarding Question 1: Note that `gse` and `gpl` do not necessarily share the same values for fields that may be identically named. That is, `gse.title` will **NOT** be the same as `gpl.title`. Thus, we should join using the `gse_gpl` file rather than trying to match on fields such as `id`. Also, when using titles to search for relevant studies, we must be careful to search for both full names and abbreviations.

```{r}
library(data.table)
gse.table <- data.table(dbGetQuery(geo_con, "SELECT gse.title, gse.gse, gse.contact FROM gse"))
gpl.table <- data.table(dbGetQuery(geo_con, "SELECT gpl.gpl, gpl.manufacturer, gpl.technology FROM gpl"))
matching.table <- data.table(dbGetQuery(geo_con, "SELECT * FROM gse_gpl"))
setkey(gse.table,gse); setkey(matching.table, gse)
##  we set keys to merge using [] operators
merged.table <- gse.table[matching.table] #note that this is a many-to-one merge
setkey(gpl.table, gpl); setkey(merged.table, gpl)
merged.table <- merge(merged.table, gpl.table, all=TRUE)
merged.table[(title %like% 'HCV' | title %like% 'Hepatitis C') & manufacturer %like% 'Illumina' & contact %like% 'Yale']
```

The entries returned by this query are the same three as were output by `SQLite`.

**This is a great HW. Too bad I'm not your partner for grading.-DEW**


